\chapter{Discussion}

My principal aims with this thesis were two-fold: first to investigate the practicality of network based analysis of complex traits and to discover general  principles that are applicable to such analyses; second to use these analyses to better understand a specific complex trait (intelligence) which we believed would be appropriate for such analysis (see section~\ref{sec:Introduction intelligence} and specifically  \ref{sec:Intelligence intro usefulness of intelligence as phenotype}).

In this chapter I will summarise the findings of the thesis for both of these. I will then discuss the strengths and limitations of this analysis and outline future work. I will conclude with some more speculative observations on this work informed by my experience in carrying out this analysis.
\section{Discussion}
\subsection{Intelligence and Educational Ability}
We find that communities associated with metabotropic glutamate receptors and ionotropic AMPA receptors are enriched for differences in intelligence and educational ability (section \ref{sec: spectral group 5 analysis}). These findings are found in discovery and replication cohorts and the findings become stronger in more recent studies that overlap with our discovery and replication cohorts but provide additional numbers. 
In the discovery and replication cohorts the enriched portions of the communities are not so much the receptors themselves but their neighbours, proteins that contribute to the macro-molecular complexes the receptors are embedded in. This finding supports the analyses and hypotheses of Grant \cite{grant2012synaptopathies} and the early work of Pocklington and co-workers \cite{pocklington2006proteomes} in community detection. 

The more recent studies (Savage and EA3)\cite{savage2018genome} \cite{lee2018gene} show enrichment both of glutamatergic receptors and their neighbours but a substantial amount of the enrichment remains to found in the neighbouring proteins. These are findings that would be missed in a standard gene set analysis approach and supports the utility of network based community gene set analysis as a method of analysing complex traits. 

For the discovery and replication cohorts the enriching proteins share few common biological themes found in Gene Ontology base gene set enrichment save for their proximity to receptor genes that are known to participate in glutamatergic neurotransmission. It is interesting to note that amongst the only findings within this group is an enrichment for neurogenesis. Hill found neurogenesis enriched in a large cohort for intelligence but that group was over 1000 genes in size, this group by contrast is 7 genes and enriches. 



This suggests that the combination of community based GSA and Gene Ontology analysis may provide a more detailed interrogation of ontology or topology defined areas of enrichment in complex traits. 


\subsection{Vertex statistic and intelligence}

No vertex statistic was associated with either intelligence or educational attainment on GWAS. That is to say there was no linear correlation between these measures. This is in contradistinction to other neurological disorders which are over represented in genes with high eigenvector centrality or degree. It is not possible to rule out a complicate or non linear relationship particularly as there is a correlation between centrality measures and loss of function intolerance and an association between loss of function intolerance and intelligence and educational ability scores for synaptic genes but no direct correlation between the two. 

\subsection{from Paper}

\section{Discussion}
Our results suggest that genetic variation in the genes encoding a group of proteins forming a tightly interconnected subnetwork in the post synaptic proteome have an enriched association with differences in educational attainment and intelligence. This group contains the majority of the genes encoding glutamate receptors and heterotrimeric G protein in the post synaptic proteome. These genes are commonly found in animal models of learning and memory and would intuitively appear to have a role in human learning and intelligence. The enrichment signal is, however, found in the neighbours of these proteins.

This paper is to our knowledge the first to use GSA in association with community detection and make use of the complete data present in GWAS summary results. Ghissian recently used the Louvain algorithm in association with the STRING protein database but did not use a tissue specific network or a competitive test of GSA and the GWA data was limited to significant genes on GWA catalog. \cite{ghiassian2015disease} 

In particular we have used an interaction network specific to a cellular component implicated in the trait. Goh found that genes that were specific to tissues were more likely to be associated with diseases than others in the connectome \cite{goh2007human}  and this seems to extend to complex traits at least in the form of intelligence and educational attainment. The review by Mooney and Wilmot of gene set analysis discusses sub communities detected in protein interaction networks as possible gene sets \cite{mooney2015gene}  and gives an example of a modularity maximisation method and an annotation mining tool that incorporates module detection; however, the first does not use genetic data and the second implements an over-representation test. does not use the full amount of data from GWA and the web address given in the paper is broken. 64  This technique could be extended to other disorders or complex traits that may be mediated by synaptic function. The synapse may behave differently from other tissues in these forms of analyses, as we know it forms modular units from its proteins that are associated with function \cite{grant2012synaptopathies} , and our finding might not generalise to proteomes in non-synaptic tissue or metabolic pathways. 

We found no association between gene level GWA p values for educational attainment or intelligence and measures of vertex importance such as degree. It may be that nodes of high degree result in a severe phenotype and are not represented in population studies of complex traits. We would hypothesise that these core nodes are necessary for basic functions, and this is supported by the finding that the nodes that have a high level of importance on each of these metrics are associated with an increased measure of evolutionary constraint. 

 	Our findings extend those of Goh to complex traits and show that genes whose variants are significantly associated with differences in cognitive ability or educational attainment are not disproportionately those of high degree.\cite{goh2007human} 
 The terms discovered using disease enrichment are heterogenous and come from animal models and databases such as OMIM and var. However, their use as indicators of an enriched association with a disease is common in the previous literature on network approaches to human disease. It is therefore desirable to have an approach incorporating all the information in the summary data of GWA study in the analysis of the association of network statistics or structures with disorders or complex traits.
 
The group is composed of glutamate receptors and G protein complexes intimately related to the plasma membrane corresponding to receptor complexes and their supporting proteins. Glutamate mediated neurotransmission is required for initiation of long term potentiation, and blocking glutamate receptors in the hippocampus of mice causes failure of memory encoding. 65,66  More recently, differential expression of glutamate receptors have recently been shown to be associated with differences in problem solving in two species of wild finch. \cite{audet2018divergence}  

The glutamate receptors and heterotrimeric G proteins are not however the source of the enrichment signal and the potential connection between the GWA findings and these receptors is made possible by this analysis of communities within the graph structure.

A small world model of genetic influence networks has been argued to support an omni-genic model of traits. \cite{boyle2017expanded}  However the structure of the synaptic graph may affect the genetic architecture of complex traits, even in a undirected network with unweighted edges and a small world structure, the rate of flow of information between genes depends on network structure.

There are limitations to the current study. The set of genes found in proteomic studies of the pre and postsynaptic areas continues to increase although the \cite{heil2018systems} increase has reached a plateau. 41  If the network changes then the community allocations will change and a different network may lead to different community allocations. We have tried to limit this effect by choosing the method with the lowest proportion of ambiguously clustered pairs (PAC).

The community allocations in addition are not a globally optimal allocation with regard to the modularity, and alternative partitions with similar modularity exist. The Spin glass model was an attractive alternative clustering method but has a number of tuneable parameters making designing a study which would properly account for multiple testing challenging.\todo{change 47 to spinglass ref}47 

The clustering also does not take consideration that nodes may belong to more than one component or cluster. Probabilistic methods exist to determine the most likely community membership if vertices can belong to more than one community73 but the loss of non-overlapping gene sets may result in a great increase in the number of sets. We have also not taken into account a model where some genes are assigned to communities while others remain unassigned.

The question of whether communities represent some ground truth entity also exists \cite{peel2017ground}  and we have not used any data other than the distribution of edges in community detection but it is possible to introduce these.  \cite{newman2016structure} 



\subsection{additional discussion}

One of the issues with the project is the continuing evolution of the synaptic maps and that we can only address two of these while they are in a continuous state of evolution. Indeed the map could be said to change with the outcome of each published synaptic proteomic study. The only way to produce anything was to freeze the data at points. Hopefully the code written to optimise the analysis will make more fluid analysis possible in the future.

Further directions will include looking at the synaptic proteome over the course of time. When moving from a trait such as intelligence to polygenic disorders that have different onsets in time then if we assume these to be synaptopathies we can expect different proteins to be expressed and this to correspond to the critical periods of the emergence of the disorder. For example neurodevelopmental disorders early, schizophrenia we would expect to occur around the time of peak onset of schizophrenia. We also know that the proteome varies across the brain and we will need to take this into account.

Future analyses may also be able to incorporate data on gene expression

One limitation is that we throw away approximately 50\% of the information from SNPs because they do not correspond to protein encoding genes. A way of incorporating genetic variants known to affect genes such as eQTLs would be helpful. 

Ideally given the continuous evolution of the map and the analysis we would have an online resource like the fly brain project that updates the findings in relatively constant times and publications to advise of major changes to the map or interface. 

The genes in the region of the afected groups are subjedt to experiment one could see if they in addition to others had a role in learning.

There are reasons to believe that the synaptic proteome is somewhat different from the general proteom and indeed that its network structure is particularly important as synaptic proteins inherently form modular units with specific functions. More phylogenetically ancient proeteins were not enriched for association for intelligence but the murine and zebrafish proteomes showed greater enrichment than the consensus proteome.

The change in GWAS size in the course of the study and the rapid change in nomenclature (for example the difficulty in translating the output of VEGAS Gene symbols to contemporary entrez show some of the factors that make the reproducibility of studies of this nature over time problematic. I have tried to minimise this by depositing all of the lookup and data information used in this project at the time it was made (eg the complete entrez info, Pli etc) so that a snapshot of the crossmapping between terms is maintained.

% EPHB2	EPH receptor B2
% 6146	RPL22	ribosomal protein L22
% 23557	SNAPIN	SNAP associated protein
% 2054	STX2	syntaxin 2
% 5127	CDK16	cyclin dependent kinase 16
% 2055	CLN8	CLN8 transmembrane ER and ERGIC protein
% 6154	RPL26	ribosomal protein L26
% 339983	NAT8L	N-acetyltransferase 8 like
% 9230	RAB11B	RAB11B, member RAS oncogene family
% 11280	SCN11A	sodium voltage-gated channel alpha subunit 11
% 2066	ERBB4	erb-b2 receptor tyrosine kinase 4
% 5138	PDE2A	phosphodiesterase 2A
% 3092	HIP1	huntingtin interacting protein 1
% 5142	PDE4B	phosphodiesterase 4B
% 83992	CTTNBP2	cortactin binding protein 2
% 55327	LIN7C	lin-7 homolog C, crumbs cell polarity complex component
% 8224	SYN3	synapsin III
% 1056	CEL	carboxyl ester lipase
% 9256	TSPOAP1	TSPO associated protein 1
% 41	ASIC1	acid sensing ion channel subunit 1
% 43	ACHE	acetylcholinesterase (Cartwright blood group)
% 11311	VPS45	vacuolar protein sorting 45 homolog
% 51248	PDZD11	PDZ domain containing 11
% 11315	PARK7	Parkinsonism associated deglycase
% 2099	ESR1	estrogen receptor 1
% 9267	CYTH1	cytohesin 1
% 5173 

You also observe the phenomenon where small changes in modularity give rise to quite different graphs. Typically this is a result of the graph having a hierarchical structure and two communities a and b being vcombined. There are a few difficulties with this first it is at present very difficult to represent such a structure using the standard tools of graph visualisation and second if one is treating communuites as units that then undergo statistical tests we have to have some way of choosing the apporirate unit. In the Newman and Givan clustering algorithm the idivision stops when the optimal level of modularity has been achieved. I would sugggest that where a study would give rise to communitiesw which make clear sense in terms of an experiment such as a populatio study where there is only a small change in modularity trhat which gives rise to the mapping that has an additional correspondance to realisty should be preferred so long as the change in modularity is not too great. This therefore results in the question of what is too great. One ansewr might be to optinmise the likelihood of hte graph and the study jointly nand in order to do this one would need a stochastic block model, this would also allow one to incorporate information on other mesoscale generating processes within the graph such as core periphery strcuture.

Does not take into account also dynamic effects e.g. rules based modelling discuss this briefly and also spatial orientation within the larger synapse (functional organisation of post synaptic glutamate receptors MacGillavry

The study provides an outline for a network analysis of GWA studies of neuropsychiatric disorders with synaptic involvement. An analysis of vertex properties with gene level results, an analysis of the site of enrichment core or periphery, community enrichment analysis and topological subgroup analysis. 

Barabasi treats disease entities as homogeneous however the disease module even if it encompasses all genes involved in the disease may represent two or more distinct processes. For example with the exception of disorders where it is possible to identify a causative mutation, organism or tissue diagnosis then a disease is a constellation of physical findings and history and investigations which may arise from more than one albeit closely related source. There are examples where redefining the groups contained within a disease group have yielded better correlation with progression and prognosis. 

One issue with MAGMA GWA is the choice of a background set. In competitive testing we test enrichment against the rest of the genome rather than seeing if the set of genes themselves are associated. The PSP enriches for educational attainment and intelligence so as the sample size of the PSP approximates the size of the PSP the enrichment value will approach this as opposed to no enrichment if the group were to approach the size of the genome. 


\subsection{Choice of clustering method.} 

A wide variety of choices of clustering method were available. Some such as the newman and girvan betweenness method were prohibitively slow with large networks even using a highly parrallelised implementation. Others such as the greedy agglomerative method led to the domination of the clusterings by some large units and many small ones. The spin glass method is also attractive but it is hard to see how one can devise an experiment using it. There is at minimum one tune able parameter and this can take any integer value. This is useful in controlling the size of groups and one can achieve good functional enrichment but the question of the choice of parameter remains an open one. One could try to learn an appropriate hyperparameter but against what is ones objective function. Both CMcC and I(in the pilot study) used grid search with more precision in the case of CMcC but this leads to a combinatorial explosion of possible clusterings and this only takes into account the gamma parameter. I think that this method is much more useful for exploratory data analysis and once one has chosen a good gamma or cooling parameter perhaps based on functional classification then one can go on to ask questions of empirical population data. However there is a problem too with this, we have shown that


 often it is not the predominant ontology member in a group that is associated with the phenotype but an agglomeration of supporting proteins which often have little in common other than their close interaction with the ontology members found to be associated with a phenotype in animal studies. In pursuing a very 'pure' functional enrichment using gene ontology for example as ones objective function using methods that have tunable parameters one misses these associated proteins. 

\subsection{Community structure and choice of algorithm}

Although the summation constraint \todo{look up} may avoid the trivial modularity solution (putting all vertices in one module) it does not avoid the 'near trivial' situation such as cone community with 99\% of the nodes and another with 1\%. Pocklington used visual inspection to fine tune the initial clustering but this is difficult when we have large numbers of vertices and is a potential source of bias if we are doing inference for example using population data. Some clustering modalities such as spin glass allow control over the number of communities with the approximate influence of size but there is the problem of tunable paramaters. Other such as Louvain are heirarchical (CDM is too) but often the implementation does not allow access tro the heirarchy or one is back with the problem of choosing the division that looks right (ie the level in the height of the heurarchy) rather than maximising the modularity or likelihood with consequent problems in inference. 

The best pragmatic solution so far may be to choose a method that gives `pretty good` (with the earlier caveat about problems in looking for too pure a functional enrichment)ideally one without tubale parameters. 

In our experience of the synaptic proteome spectral clustering and Louvain perform well although the groups are rather large with Louvain. In the future especially with access to raw genotype data maximising the joint (improbability) of the population study and clustering may be effective ie we accept a small decrease in modularity if it is associated with a partition that makes sense of the study but then there is the question of how small the change in modularity can be. 



We need a way in which we can take into account that the pre and post synaptic areas are integrated eg some processes cross them and form cross synaptic modules and also that some genes are present in both compartments. It seems that the isolated post synaptic component is enriched for educational attainment but a number of the most significant genes are those seen in the PSP.

In considering the difference between disease modules and topological groups we must consider two possibilities. First for a complex trait we are looking at a topological area where genetic variation occurs across the population which may lead to a convergent phenotype. This is different from looking at the effects of the perturbation of an area of the graph on neighbouring or more distant nodes for example gene expression in a loss of function mutation. OMIM and Var contain both, genes implicated in disease that are up-regulated or found to have a causative role and those where there are specific germ line changes that predispose to the disease. The modular structure of the synapse and the effects of synaptic protein structure on function make it more likely that differences in neighbouring proteins will impact the overall function of a topological unit or the synapse than in for example a biological pathway where constituents are more separated in time than space and limiting factors may include quantity of reactants. We see in the study of core genes the effects that variations in these have in multi-system disorders, incompatibility with viability or disorders of cell cycle (which is an ancient system conserved from yeast)

\section{For discussion but not sure what to say yet}

For the presynaptic proteome education phenotypes enrich but the intelligence ones in general do not. 

\section{Further discussion}
Lack of different intelligence phenotypes e.g. fluid etc in large scale genomic studies compared with Hill et al. where the finding was for a single phenotype
\section{Further research}
Missing heritability and the impact of exome data

Make it easy to use like FUMA

How findings of GWAS extend to more rare diseases give example of drug induced psychosis or high intelligence